Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-05-18 bbbq_article.tex
Supmat fixed
Add dataset strength calculation

\begin{table}
  \input{bbbq_1/bbbq_1_percentages.latex}
  \caption{
    Percentage of MHC-I epitopes overlapping with transmembrane helix.
    \richel{This is simulated data}
  }
  \label{table:bbbq_1_percentages}
\end{table}

\begin{table}
  \input{bbbq_1/bbbq_1_stats_covid.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-I
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_1_stats_covid}
\end{table}

\begin{table}
  \input{bbbq_1/bbbq_1_stats_myco.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-I
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_1_stats_myco}
\end{table}

@@

\begin{table}
  \input{bbbq_2/bbbq_2_percentages.latex}
  \caption{
    Percentage of MHC-II epitopes overlapping with transmembrane helix.
    \richel{This is simulated data}
  }
  \label{table:bbbq_2_percentages}
\end{table}

\begin{table}
  \input{bbbq_2/bbbq_2_stats_covid.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-II
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_2_stats_covid}
\end{table}

\begin{table}
  \input{bbbq_2/bbbq_2_stats_myco.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-II
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_2_stats_myco}
\end{table}

@@

\begin{table}
  \input{mhc2_haplotypes.latex}
  \caption{
    Percentage of MHC-II haplotypes, from \cite{greenbaum2011functional}
    \richel{
      This is just a reminder, instead of new research. 
      This table be deleted in the future.
    }
  }
  \label{table:mhc2_haplotypes}
\end{table}

>>>>

\begin{table}[!htbp]
  \input{bbbq_1/bbbq_1_percentages.latex}
  \caption{
    Percentage of MHC-I epitopes overlapping with transmembrane helix.
    \richel{This is simulated data}
  }
  \label{table:bbbq_1_percentages}
\end{table}

\begin{table}[!htbp]
  \input{bbbq_1/bbbq_1_stats_covid.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-I
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_1_stats_covid}
\end{table}

\begin{table}[!htbp]
  \input{bbbq_1/bbbq_1_stats_myco.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-I
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_1_stats_myco}
\end{table}

@@

\begin{table}[!htbp]
  \input{bbbq_2/bbbq_2_percentages.latex}
  \caption{
    Percentage of MHC-II epitopes overlapping with transmembrane helix.
    \richel{This is simulated data}
  }
  \label{table:bbbq_2_percentages}
\end{table}

\begin{table}[!htbp]
  \input{bbbq_2/bbbq_2_stats_covid.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-II
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_2_stats_covid}
\end{table}

\begin{table}[!htbp]
  \input{bbbq_2/bbbq_2_stats_myco.latex}
  \caption{
    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-II
    \richel{Done on the simulated data}
  }
  \label{table:bbbq_2_stats_myco}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Strength of COVID-19 dataset}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Our evolutionary experiment depends on the strength of the dataset.
Here we do a back-on-the-envelope calculation to assess if the dataset
is strong enough.

\begin{itemize}
	\item Genome size: $\approx 3 \cdot 10^4$ bases
	\item Evolution rate: $\approx 10^{-3}$ per nucleotide per year (from \cite{bar2020sars})
	\item Evolution rate per genome: $\approx 26$ per year (GISAID)
	\item Evolution rate per nucleotide: $0.00088$ per nucleotide per year as $\frac{26}{\cdot 10^4}$ per nucleotide year
	\item Mutation rate: $\approx 10^{-6}$ per nucleotide per cycle
	\item Proteome size: 9701 (from reference proteome)
	\item Size of proteome compared to genome: $\frac{9701}{3 \cdot 10^4} \approx \frac{1}{3}$ \richel{no/little non-coding DNA indeed}
	\item Fraction of mutations that change AA: $\frac{20}{64}$ 
	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
	\item TMHs per proteome: 21 (from TMHMM)
	\item AAs per TMHs: 23 (from Bianchi et al.)
	\item AAs per proteome that is TMH: $21 \cdot 23 = 483$
	\item Mutations in TMH per proteome year: $\approx 0.444$ (as $483 \cdot 0.00092$)
	\item Number of sequenced proteomes: 5600 (GISAID, genomes)
	\item Pandemic duration: $\approx \frac{1}{3}$ year
	\item Expected number of observed TMH mutations: $\approx 800$ (as $0.444 \cdot 5600 \cdot \frac{1}{3}$)
	\item Chance that a mutation may lose/acquire epitope binding: $\frac{1}{100}$ \richel{Just a guess, based on weak selection}
	\item Expected number of observed TMH mutations that cause a transition: $\approx 8$ (as $800 \cdot \frac{1}{100}$)
\end{itemize}

Conclusion: weak, but a good

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Strength of Influenza A}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Our evolutionary experiment depends on the strength of the dataset.
Here we do a back-on-the-envelope calculation to assess if the dataset
is strong enough.

\begin{itemize}
	\item Genome size: $\approx 1.3 \cdot 10^4$ bases
	\item Relative genome size compared to COVID-19: $\approx 0.4$ (as $\frac{1.3 \cdot 10^4}{3.0 \cdot 10^4}$)
	\item Full genome sets: $\approx 3.2 \cdot 10^4$ (Influenza A)
	\item Relative number of genome sets compared to COVID-19: $\approx 6.4$ as much (as $\frac{\approx 32 \cdot 10^3}{5 \cdot 10^3}$)
	\item Relative time to accumulate mutations: $\approx 12$ times longer (average age $\approx$ 4 year, before 2015, already 8k sequences)
	\item Relative chance of observing transitions: $\approx 30$ (as $0.4 \cdot 6.4 \cdot 12$)
	\item Expected number of observed TMH mutations in COVID-19 that cause a transition: $\approx 8$
	\item Expected number of observed TMH mutations that cause a transition: $\approx 240$ (as $8 \cdot 30$)
\end{itemize}

Conclusion: seems strong enough to us.

@@

\begin{table}[!htbp]
  \input{mhc2_haplotypes.latex}
  \caption{
    Percentage of MHC-II haplotypes, from \cite{greenbaum2011functional}
    \richel{
      This is just a reminder, instead of new research. 
      This table be deleted in the future.
    }
  }
  \label{table:mhc2_haplotypes}
\end{table}
